Computing Entropies With Nested Sampling

### 2017-06-19 paper/ms.tex
Fixed equation

The sequence of $\ln X$ values produced by a Nested Sampling run can be
considered as a Poisson process with rate equal to $N$, the number of
particles. For example, when $N=1$, the distribution for $\ln X_{i+1}$
given $\ln X_i$ is a shifted (and reflected) exponential with rate 1.
\citet{Walter2015} showed how this view of the NS sequence of points
can be used to construct a
version of NS that produces unbiased (in the frequentist sense)
estimates of the evidence.
However, it can also be used to construct an unbiased estimator of
a log-probability, which is more relevant to information theoretic
quantities.

Consider a particular likelihood value $\ell$ whose corresponding $X$-value
we would like to know. Since the $\ln X_i$ values have the same distribution
as the arrival times of a Poisson process with rate $N$, the probability
distribution for the number of points in an interval of length $t$ is
Poisson with expected value $\lambda t$. Therefore the expected number of
NS discarded points with likelihood below $\ell$ is $-\ln X(\ell)$:
\begin{align}
\big< n(L(\x_i) \leq \ell) \big> &= -\ln X(\ell).
\end{align}
We can therefore take $n(L(\x_i) \leq \ell)$, the number of points in the
sequence with likelihood below $\ell$, as an unbiased estimator of
$-\ln X(\ell)$.

@@

The above insight, that the number of NS discarded points with likelihood
below $\ell$ has expected value $-\ln X(\ell)$, is the basis of the algorithm.
If we wish to measure a log-probability $\ln X(\ell)$, we can use NS to do it.
See Algorithm~\ref{alg:algorithm} for a step-by-step description of the
algorithm. The algorithm is written in terms of an unknown quantity $\x$
whose entropy is required. In applications, $\x$ may be parameters,
a subset of the parameters, data, some function of the data,
parameters and data jointly, or whatever.

>>>>

The sequence of $-\ln X$ values produced by a Nested Sampling run can be
considered as a Poisson process with rate equal to $N$, the number of
particles. For example, when $N=1$, the distribution for $-\ln X_{i+1}$
given $-\ln X_i$ is a shifted (and reflected) exponential with rate 1.
\citet{Walter2015} showed how this view of the NS sequence of points
can be used to construct a
version of NS that produces unbiased (in the frequentist sense)
estimates of the evidence.
However, it can also be used to construct an unbiased estimator of
a log-probability, which is more relevant to information theoretic
quantities.

Consider a particular likelihood value $\ell$ whose corresponding $X$-value
we would like to know. Since the $-\ln X_i$ values have the same distribution
as the arrival times of a Poisson process with rate $N$, the probability
distribution for the number of points in an interval of length $t$ is
Poisson with expected value $\lambda t$. Therefore the expected number of
NS discarded points with likelihood below $\ell$ is $-N\ln X(\ell)$:
\begin{align}
\big< n(L(\x_i) \leq \ell) \big> &= -N\ln X(\ell).
\end{align}
We can therefore take $n(L(\x_i) \leq \ell)/N$, the number of points in the
sequence with likelihood below $\ell$ divided by the number of NS particles,
as an unbiased estimator of $-\ln X(\ell)$.

@@

The above insight, that the number of NS discarded points with likelihood
below $\ell$ has expected value $-N\ln X(\ell)$, is the basis of the algorithm.
If we wish to measure a log-probability $-\ln X(\ell)$, we can use NS to do it.
See Algorithm~\ref{alg:algorithm} for a step-by-step description of the
algorithm. The algorithm is written in terms of an unknown quantity $\x$
whose entropy is required. In applications, $\x$ may be parameters,
a subset of the parameters, data, some function of the data,
parameters and data jointly, or whatever.
